%!TEX root = focs.tex


\section{Games with constant payoff [simplified version]}
\label{sec-const_rew}

The first version of the game we analyse is when the payoff function $r_p(q)$ pays each block in the blockchain the same amount $c$. While this does not reflect the current reality of the Bitcoin protocol, this simplification serves as a good baseline for future results, and it establishes the main techniques we will have to utilize. Furthermore, the results we obtain here could serve as a good recommendation of how the Bitcoin protocol can be modified in order to enforce fair behaviour of the miners when the block reward becomes insignificant.

When considering the constant reward $c$ for each block, $r_p(q)$ will equal $c$ times the number of blocks owned by $p$ in $\bchain(q)$, when the latter is defined. On the other hand, when $\bchain(q)$ is not defined it might seem tempting to simply define $r_p(q) = 0$. However, even if there is more than one longest branch from the root of $q$ to its leaves, it might be the case that all such branches share a common subpath. In fact, this often happens in the Bitcoin's network, when two blocks were mined on top of the same block (with a small time delay). While in this situation the blockchain is not defined, the miners know that they will at lest be able to collect their reward on the portion of the body of knowledge these two branches agree on. Figure \ref{fig-simple-fork} illustrates this situation. In order to model this scenario we will have to introduce some more notation. 

\begin{figure}
\begin{center}
\begin{tikzpicture}[->,>=stealth',auto,thick, scale = 1.0,state/.style={circle,inner sep=2pt}]

    % The graph
	\node [state] at (0,0) (R) {$\varepsilon$};
	\node [state] at (1.5,0) (1) {$1$};
	\node [state] at (3,0) (10) {$10$};
	\node [state] at (4.5,0) (100) {$100$};
	\node [state] at (6,0) (1001) {$1001$};
	\node [state] at (8,0.75) (10011) {$10011$};
	\node [state] at (8,-0.75) (10010) {$10010$};	
	
	% Graph edges
	\path[->]
	(R) edge (1)
	(1) edge (10)
	(10) edge (100)
	(100) edge (1001)
	(1001) edge (10011)
	(1001) edge (10010)
	;  	
	


\end{tikzpicture} 
\end{center}
\label{fig-simple-fork}
\caption{Although two branches are contesting to become a blockchain, the blocks up to 1001 will contribute to the reward in each case.}
\end{figure}


Recall that a block $b$ is a string over the alphabet $\bP$, and we use notation $|b|$ for the length of $b$ as a string. Moreover, given blocks $b_1, b_2$, we use notation $b_1 \preceq b_2$ to indicate that $b_1$ is a prefix of $b_2$  when considered as strings. Then we define: 
\begin{eqnarray*}
\longest(q) & = & \{ b \in q \mid \text{for every } b' \in q: |b'| \leq |b|\}\\
\meet(q) & = & {\displaystyle \max_{\preceq} \ \{b \in q \mid \text{for every } b' \in \longest(q): b \preceq b'\}}
\end{eqnarray*}
Intuitively, $\longest(q)$ contains the leaves of all branches in the state $q$ that are currently competing for the blockchain, and $\meet(q)$ corresponds to the last block for which all these branches agree on. For instance, if $q$ is the state from Figure \ref{fig-simple-fork}, then we have that $\longest(q)=\{10011,10010\}$, and $\meet(q)=1001$. Notice that $\meet(q)$ is well defined as $\preceq$ is a linear order on the finite and non-empty set $\{b \in q \mid \text{for every } b' \in \longest(q): b \preceq b'\}$. Also notice that $\meet(q)=\bchain(q)$, whenever $\bchain(q)$ is defined.

As mentioned above, the pay-off function will reward a player for the blocks she owns in the path from the genesis block to $\meet(q)$. Thus, to define the pay-off, we need to identify who is the owner of each one of these blocks, which is done by considering the function $\chi_p$, for each $p \in \bP$. More precisely, given $b \in \bB$ and $i \in \{1, \ldots, |b|\}$, we have that:
%First, for a player $p$, define the indicator function $\chi_p : \bB \to \{0,1\}$ as follows:
\begin{eqnarray*}
\chi_p(b,i) & = & 
\begin{cases}
1 & \text{the } i\text{-th symbol in } b \text{ is } p\\
0 & \text{otherwise}
\end{cases}
\end{eqnarray*}

We can finally define the payoff function we consider in this section, which we call the \textbf{constant reward}. For a player $p$, we define it as 
\begin{eqnarray*}
r_p(q) & = & 
{\displaystyle c \cdot \sum_{i=1}^{|\meet(q)|} \chi_p(\meet(q),i)} 
\end{eqnarray*}
where $c$ is a positive real number. As mentioned above, this function is well defined, since $\meet(q)$ is always defined, and equals $\bchain(q)$ when the latter is defined for the state $q$.
%Here we use notation $b[i]$ for the $i$-th symbol in $b$, where $i \in \{1, \ldots, |b|\}$. 
% and $d \in \mathbb{N}$. Here the number $d$ is the amount of confirmations needed to spend the block (6 in the case of Bitcoin).
 
\subsection{The default strategy}

Let us start with analysing the most obvious strategy for all the players: regardless of what everyone else does, keep mining on the blockchain. We call this 
the \emph{default} strategy, as is it reflects the desired behaviour of the miners participating in the Bitcoin network. For a player $p$, let us denote this strategy 
by $\df_p$, and consider the combined strategy $\df = (\df_1,\dots,\df_p)$. Notice that under this strategy, each state $q$ consists of a single path from the genesis block $\varepsilon$ to the final block in $\bchain(q)$.

We can now easily calculate the utility of player $p$ under $\df$. Intuitively, a player $p$ will receive a fraction $h_p$ of the next block that is being placed in the blockchain, corresponding to her hash power. Therefore, at stage $i$ of the mining game, $i$ blocks will be placed in the blockchain defined by the game, and the expected amount of blocks owned by the player $p$ will be $h_p\cdot i$. This means that the total utility for player $p$ amounts to $$u_p(\bs \mid q_0) = c\cdot h_p \cdot \sum_{i=0}^{\infty}i \cdot \beta^{i}.$$ %(recall that $q_0$ is the genesis block). 

The question the is: can any player do better? As we will show, the answer is no if we assume that the rest of the players behave according to $\df$. In the remainder of the section we 
will prove that $\df$ is a $\beta$-discounted stationary equilibrium, under a very mild assumption on the set of strategies available to each player. 
To prove this fact we need to show that the utility of player $p$ under any strategy $(\df_{-p},s)$ is not higher than the utility under $\df$. Note that we only need to focus on 
games for two players, as all the players distinct from $p$ can be combined into a single player with their combined hash power. 

The key lemma we use to prove this result (and several others) is the fact that, in our game, the optimal strategies for a player $p$ are determined only by the portion 
of blocks that appear after the last block owned by $p$ in the blockchain. In the following subsection we sketch the proof of these results. A detailed proof, together with the full definition of default strategies, can be found in Appendix \ref{app-const}.

\subsection{Equilibria in two-player games} 

In this section we will consider games with two players, that is, $\bP = \{0,1\}$. Our main objective is to show that the default strategy is indeed an equilibrium. For this, we will slightly restrict the space of strategies that the player use, and concentrate on the so called {\em greedy} strategies. Intuitively, under greedy strategies, the players refrain from forking on top of blocks that appear before their latest block when there is no blockchain, or their latest block in the blockchain, when the latter is defined. Greedy strategies can be formally defined as follows.



Given a player $p \in \{0,1\}$ and a state $q \in \bQ$, define:
%\begin{multline*}
%\longest(q,p) \ = \ \{ b \in q \mid b = \varepsilon \text{ or } \owner(b) = p,\\
%\text{ and for every } b' \in q \text{ such that } \owner(b') = p : |b'| \leq |b|\}.
%\end{multline*}
$$
\longest(q,p) \ =
\begin{cases}
\{\varepsilon\}, \text{ if } \forall b\in q:\ \owner(b)\neq p\\
\{b\in q \ \mid \ \owner(b)=p \text{ and } \forall b'\in q:\ \owner(b')=p \text{ implies } |b'|\leq |b|\}, \text{ otherwise}
\end{cases}
$$
%Notice that $\varepsilon \in \longest(p,q)$ if and only if there is no $b \in q$ such that $\owner(b) = p$. 
Moreover, define $\length(q,p)$ as the length of an arbitrary string in $\longest(q,p)$ (notice that all of them have the same length).
\begin{mydef}\label{def-greedy}
Given $p \in \{0,1\}$, $b \in \bB$ and $q \in \bQ$,  an action $\mine(p,b,q)$ is {\em greedy} if $\mine(p,b,q)$ is a valid action and $\length(q,p) \leq |b|$.
%
Moreover, a strategy $\bs = (s_0, s_1)$ is {\em greedy} if for every $p \in \{0,1\}$ and  $q \in \bQ$ such that $\pr^{\bs}(q \mid \varepsilon) > 0$, it holds that $s_p(q)$ is a greedy action.
\end{mydef}
From now on, we only consider greedy strategies. 
%
%Under greedy strategies, players refrain to fork on top of blocks that appear before their latest block, or their latest block in the blockchain. 
The consequence of this is that every state in a two-player game under greedy strategies cannot have more than two paths contesting for the blockchain (see Lemma \ref{lem-length-greedy} in Appendix \ref{app-const}). %This is captured b the following technical lemma: 
%
%\begin{mylem}\label{lem-length-greedy}
%Let $\bs$ be a greedy strategy. Then for every $q \in \bQ$ such that $\pr^{\bs}(q \mid \varepsilon) > 0$, the following conditions hold:
%\begin{enumerate}
%\item For every $p \in \{0,1\}$: $|\longest(q,p)| = 1$ 
%
%\item $1 \leq |\longest(q)| \leq 2$
%
%\item If $|\longest(q)| = 2$, then $\longest(q) = \longest(q,0) \cup \longest(q,1)$
%\end{enumerate}
%\end{mylem}
Similarly, we can also show that if only greedy strategies are considered, an optimal strategy for player $p$ can only differentiate the portion of a state $q$ that begins after $\meet(q,p)$. That is, for any two equilibria $s=(s_1,s_2)$ and $s'=(s_1',s_2')$, and any two states $q,q'$ such that the tree rooted at $\meet(q,p)$ equals the tree tooted at $\meet(q',p)$, it holds that $s_p(q)=s_p'(q')$ (see Lemma \ref{lem-optimal} below).




%\paragraph{Longest blocks and optimal strategies.}
%For a state $q$ and a block $b \in q$, let us denote by $\subbody(q,b)$ the state
%given by $\{u \mid b\cdot u$ is a block in $q\}$, that is, the subtree of $q$ rooted at $b$, but in which $b$ is renamed 
%$\epsilon$ and all its descendants are renamed accordingly. 
%
%%Furthermore, let us denote by $\meet(q,p)$ the greatest block in the set $\{b \in q \mid b$ is a prefix all nodes in $\longest(q)\}$, the greatest common block owned by $p$ that is a prefix of all 
%%blocks in $\longest(q)$. 
%
%The following Lemma tells us that an optimal strategy for player $p$ can only differentiate the portion of 
%a state that goes after $\meet(q,p)$: 
%
%
%\begin{mylem}
%Let $s = (s_1,s_2)$ be a $\beta$ discounted stationary equilibrium in an infinite mining game with two players. 
%Then there is a $\beta$ discounted stationary equilibrium such that $u_p(s \mid q_0) = u_p(s' \mid q_0)$ for 
%any player $p$ and for every pair $q$ and $q'$ of 
%bodies of knowledge in which $\subbody(q,\meet(q,p)) = \subbody(q',\meet(q',p))$ we have that 
%$s_p(q) = s_p(q')$. 
%\end{mylem}
%
%\begin{proof}
%\end{proof}

Armed with these two observations, we can now obtain our first main result:

\begin{mythm}\label{thm-conts_equlibria}
For any $0 \leq \beta \leq 1$, the strategy $\df$ is a $\beta$-discounted stationary equilibrium under greedy strategies. 
\end{mythm} 

\begin{proof}[Proof sketch]
recall two players, greedy, use lemma above. 
\end{proof}

We already mentioned that constant block rewards do not faithfully model reality, since in the Bitcoin protocol the reward decreases every 200.000 blocks or so. However, we would like to argue that Theorem \ref{thm-conts_equlibria} can serve as a good recommendation on how to enforce good behaviour on miners (assuming they will use the utility function as an indicator of their monetary gain), at the moment block rewards become insignificant. More precisely, if block rewards are insignificant, and the transaction fees dictate the miners' pay-off, the protocol could place a (constant) fee limit on newly created blocks. Assuming that the volume of transactions is high, the blocks would regularly achieve the maximal reward, thus making the block reward constant. Theorem \ref{thm-conts_equlibria} then tells the miners that their best strategy is to mine on top of the existing blockchain, as this will maximize their utility in the long run.