%!TEX root = focs.tex


\section{Off}

Section under construction, mostly to share my preliminary results with you.

\iffalse

\begin{mydef}
	Let $p_1,p_2$ two players, we say that a state $q$ is \textbf{steady} if and only if : 
	\begin{align*}
		\forall \pi_1,\pi_2 \in \cT(q), \pi_1 \not \subseteq \pi_2 \land \pi_2 \not \subseteq \pi_1 \Rightarrow & \pi_1 \cap \bchain(q) \neq \pi_1 \cap \bchain(q) \\
		& \land \left[
		\begin{array}{ll}
		\forall b \in (\pi_1 \setminus \bchain(q)), owner(b) = 1 \lor owner(b) = 2 \\
		\forall b \in (\pi_2 \setminus \bchain(q)), owner(b) = 1 \lor owner(b) = 2 \\
		\end{array}\right] \\ 
	\end{align*}
\end{mydef}

\begin{mydef}
	Let $p_1,p_2$ two players, $q$ a state, let $\pi_0$ the longest path of $q$ such that $owner(\pi_0[|\pi_0|]) = 1$.
	We say that a strategy $s$ for player 1 is \textbf{consistent} if and only if : 
	$$
	 s(q) = \left\{
	 \begin{array}{ll}
	 \mine(p_1,\last(\bchain(q)),q) & \textit{ iff } \pi_0 = \bchain(q). \\
	 \mine(p_1,b,q) \textit{ such that } b\in \{\bchain(q)\setminus \pi_0\} \cup \{\last(\pi_0)\} & \textit{ otherwise.} \\
	 \end{array}\right. \\ 
	$$
\end{mydef}

\begin{myprop}
	\label{steadyprop}
	For every steady state $q$, a combined strategy $\bs$ is a $\beta$- discounted nash equilibrium if and only if $\forall s \in \bs$, $s$ is consistent.
\end{myprop}

\begin{proof}
	\begin{mylem}
		Let $\bs$ a $\beta$- discounted nash equilibrium and $q$ a state such that $s_1(q) = \mine(p_1,\last(\pi),q)$ then $$\forall q', q \subset q' \land s_1(q') =  \mine(p_1,\last(\pi'),q') \Rightarrow |\pi'| > |\pi|$$
	\end{mylem}
	Assuming lemma one, it is easy to demonstrate prop \ref{steadyprop}. Indeed by absurd if it is not true for any $q$ steady it is not true either for "base case scenario".
	\etienne{I am still working on the lemma proof, so the result is useless so far ... I am getting familiar with your math to see if i can get some idea from it.}
\end{proof}

\begin{myprop}
	\label{consprop}
	For every steady state $q$, and $\beta$- discounted nash equilibrium $\bs$, we have $\forall s \in \bs, s(q)(q) \textit{ is steady}$.
\end{myprop}
\begin{proof}
	It is immediate by proposition \ref{steadyprop} and the definition of a consistent strategy.
\end{proof}

\begin{myprop}
	Every attainable state in full disclosure two players game under a $\beta$- discounted nash equilibrium is steady.
\end{myprop}
\begin{proof}
	By induction, we know that $\epsilon$ is steady and with property \ref{consprop} we can conclude. 
\end{proof}


\fi
\begin{mylem}
	Let $s = (s_1,s_2)$ be a $\beta$ discounted stationary equilibrium in an infinite mining game with two players. 
	Then there is a $\beta$ discounted stationary equilibrium such that $u_p(s \mid q_0) = u_p(s' \mid q_0)$ for 
	any player $p$ and for every pair $q$ and $q'$ of 
	bodies of knowledge in which $\subbody(q,\mid(q,p)) = \subbody(q',\mid(q',p))$ we have that 
	$s_p(q) = s_p(q')$. 
\end{mylem}

\begin{proof}
	Let $\bs = (s_1,s_2)$ a combined greedy strategy. 
	Let $q$ a state, such that $\pr^{\bs}(q \mid q_0) \neq 0$. 
	Then either $\subbody(q,\mid(q,p_1)) \subseteq \subbody(q,\mid(q,p_2))$ or $\subbody(q,\mid(q,p_2)) \subseteq \subbody(q,\mid(q,p_1))$ by definition of $\mid(q,p)$.
	
	Then we  can define $q_{max} = \underset{p \in \{p_1,p_2\}}{max}\subbody(q,\mid(q,p))$
	
	We also define $b^q_{min} = max\{b \mid b \textit{ strict prefix} \mid(q,p_1) \land b \textit{ strict prefix} \mid(q,p_2) \}$
	
	Note that $longest(q) = b^q_{min}.longest(q_{max})$
	
	\begin{mylem}
	Let $q'$ a state, such that $\pr^{\bs}(q' \mid q) \neq 0$. 
	Then  $\forall b \in longest(q'), b^q_{min} \textit{ prefix of } b$
	\end{mylem}
	\begin{proof}
		by induction : 
		We just have to prove : 
		\begin{itemize}
		\item $\forall b \in longest(s_1(q)), b^q_{min} \textit{ prefix of } b $
		\item $\forall b \in longest(s_2(q)), b^q_{min} \textit{ prefix of } b $
		\end{itemize}
	Which is trivial as $s_1$ and $s_2$ are greedy.
	\end{proof}

Especially the previous lemma means that $\forall q'$ such that $\pr^{\bs}(q' \mid q) \neq 0$ and $bc(q')$ is defined then $b^q_{min}$ is a prefix of $bc(q')$


\etienne{Here it is not true yet because of the definition of reward i will fix it later, but i need to move on for now and the logic behind is the same it just simplified notation and cases. (I dont care about d-delayed and stuff)}

\begin{mylem}
	Let $q'$ a state, such that $\pr^{\bs}(q' \mid q) \neq 0$, let $b$ such that $bc(q') = b^q_{min}.b$ ($b$ exists by previous lemma.) then there exits a function $f_{b^q_{min}}$ such that :
	Then  $$r_p(q') = r_p(b^q_{min}) + f_{b^q_{min}}(r_p(b))$$
\end{mylem}
\begin{proof}
	I have to consider every reward for instance:  $\alpha$ discounted reward: 
	$r_p(q') = r_p(b^q_{min}) + \alpha^{|b^q_{min}|}r_p(b)$
\end{proof}

Moreover
\begin{eqnarray*}
	u_p(\bs \mid q) & = & \sum_{i=0}^{\infty}\beta^{i} \cdot  \bigg(\sum_{\substack{q' \in \bQ \,: \\ q \subseteq q' \text{ {\rm and} } |q'| - |q| = i}} r_p(q') \cdot 
	\pr^{\bs}(q' \mid q)\bigg)
\end{eqnarray*}

By the two previous lemma (as either $\pr^{\bs}(q' \mid q) = 0$ or true) 

\begin{eqnarray*}
	u_p(\bs \mid q) & = & \sum_{i=0}^{\infty}\beta^{i} \cdot  \bigg(\sum_{\substack{q' \in \bQ \,: \\ q \subseteq q' \text{ {\rm and} } |q'| - |q| = i}} [r_p(b^q_{min}) + f_{b^q_{min}}(r_p(b))] \cdot 
	\pr^{\bs}(q' \mid q)\bigg) \\ 
	& = & \sum_{i=0}^{\infty}\beta^{i} \cdot  r_p(b^q_{min})  \\ 
	&&+ \sum_{i=0}^{\infty}\beta^{i} \cdot  \bigg(\sum_{\substack{q' \in \bQ \,: \\ q \subseteq q' \text{ {\rm and} } |q'| - |q| = i}}  f_{b^q_{min}}(r_p(b)) \cdot 
	\pr^{\bs}(q' \mid q)\bigg) \\ 
\end{eqnarray*}

Let $\bs' = (s_1',s_2')$ a combined greedy strategy. 
Then 

\begin{eqnarray*}
	u_p(\bs \mid q) - u_p(\bs' \mid q)& = & \sum_{i=0}^{\infty}\beta^{i} \cdot  r_p(b^q_{min})  \\ 
	&&+ \sum_{i=0}^{\infty}\beta^{i} \cdot  \bigg(\sum_{\substack{q' \in \bQ \,: \\ q \subseteq q' \text{ {\rm and} } |q'| - |q| = i}}  f_{b^q_{min}}(r_p(b)) \cdot 
	\pr^{\bs}(q' \mid q)\bigg) \\ 
	&& -  \sum_{i=0}^{\infty}\beta^{i} \cdot  r_p(b^q_{min})  \\ 
	&& - \sum_{i=0}^{\infty}\beta^{i} \cdot  \bigg(\sum_{\substack{q' \in \bQ \,: \\ q \subseteq q' \text{ {\rm and} } |q'| - |q| = i}}  f_{b^q_{min}}(r_p(b)) \cdot 
	\pr^{\bs'}(q' \mid q)\bigg) \\ 	
	& = & \sum_{i=0}^{\infty}\beta^{i} \cdot  \bigg(\sum_{\substack{q' \in \bQ \,: \\ q \subseteq q' \text{ {\rm and} } |q'| - |q| = i}} f_{b^q_{min}}(r_p(b)) \cdot 
	[\pr^{\bs}(q' \mid q)-\pr^{\bs'}(q' \mid q)]\bigg) \\
\end{eqnarray*}

And it is done ! (As the difference of utility doesn't rely on the state before $b^q_{min}$ and by definition we have : 
$q'_{max} = q_{max} \Rightarrow s_p(q) = s_p(q')$. Which is not exactly the lemma but by application of greedy actions rule we obtain it.

\end{proof}

\etienne{What i am currently working on : }
\begin{myprop}
	Let $\bs = (s_1,s_2)$ a $\beta$ discounted nash equilibrium then $\bs$ is greedy.
\end{myprop}
\begin{proof}
	\begin{mylem}
		Let $\bs = (s_1,s_2)$ a $\beta$ discounted nash equilibrium such that $s_1$ is greedy then $s_2$ is greedy.
	\end{mylem}
\end{proof}





